```{r}
library(tidyverse)
```
```{r}
library(dplyr)
```
```{r}
library(ggplot2)
```
```{r}
library(corrgram)
```


```{r}
## n_distinct(PGA$Player_initial_last)
```

```{r}
PGA<-read_csv('ASA All PGA Raw Data - Tourn Level.csv')
```

##PGA



Here I am finding some average summarys statistics from the PGA dataset to get a 
feel for what the norm of these stats are. They will range from higher to lower 
depending on the level of golfers I am looking at. 
```{r}
(summary <- PGA %>% 
  filter(!is.na(made_cut)) %>% 
  group_by(made_cut) %>% 
  summarise(avg_strokes = mean(strokes, na.rm = TRUE),
            avg_sg_put = mean(sg_putt, na.rm = TRUE),
            avg_sg_app = mean(sg_app, na.rm = TRUE),
            avg_sg_ott = mean(sg_ott, na.rm = TRUE),
            avg_sg_t2g = mean(sg_t2g, na.rm = TRUE)))
```
##top 25
Filtering the top 25 players in the total pga dataset. 
```{r}
(PGA_top25<- PGA %>% 
  filter(!is.na(pos)) %>% 
  arrange(pos) %>% 
  slice(1:25) %>% 
  select(Player_initial_last, pos, strokes, sg_putt, sg_app, sg_ott, sg_total, sg_arg, sg_t2g))
```

Creating new variable that sums up specifically the short game, which I think is a 
contributing factor to being good at the game. I will look into it further. 
```{r}
(PGA_top25 <- PGA_top25%>% 
  mutate(short_game = sg_putt + sg_arg))
```



## top 10
Filtering through the data set to find the top 10 players by position, also choosing 
key stats so we can compare them.
```{r}
PGA_top10<- PGA %>% 
  filter(!is.na(pos)) %>% 
  arrange(pos) %>% 
  slice(1:10) %>% 
  dplyr::select(Player_initial_last, pos, strokes, sg_putt, sg_app, sg_ott, sg_total, sg_arg)
```

Adding the variable short game to PGA_top 10 .
```{r}
(PGA_top10 <- PGA_top10%>% 
  mutate(short_game = sg_putt + sg_arg))
```

```{r}
ggplot(PGA_top10, aes(x = reorder(Player_initial_last, -short_game), y = short_game, color = Player_initial_last)) +
  geom_point(size = 4) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```
This plot demonstrates that players with higher short-game totals which are positive strokes gained 
were able to help them save strokes on and around the greens contributing to a lower finishing place.
0.00 is avg, +1.5 is better than avg, and +2.0 is elite
The higher the score the more strokes youre saving per round.


Filtering through the data set to find the worst ranked pga golfers.
```{r}
PGA_bottom25 <- PGA %>%
  filter(!is.na(pos)) %>%
  arrange(desc(pos)) %>%
  slice(1:25) %>%
  dplyr::select(Player_initial_last, pos, strokes, sg_putt, sg_arg, sg_app, sg_ott, sg_total, sg_t2g) %>%
  mutate(short_game = sg_putt + sg_arg)
```


```{r}
ggplot(PGA_bottom25, aes(x = reorder(Player_initial_last, -short_game), y = short_game, color = Player_initial_last)) +
  geom_point(size = 4)+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```
Here is the contrary showing the worst players in the database and how their short-game
totals are mostly negative causing them to have more strokes per hole and place lower.
Here we can see that a lot of these guys are in the negatives causing them to gain
strokes per round. There are a few guys above 0 meaning they have an average stat
but their short game could be bad which would make sense on why they place higher. 


```{r}
ggplot(PGA_top25, aes(x = reorder(Player_initial_last, -sg_t2g), y = sg_t2g, fill = Player_initial_last)) +
  geom_col(show.legend = FALSE) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```
Here we can see the top 25 avg shots to the green.


```{r}
ggplot(PGA_bottom25, aes(x = reorder(Player_initial_last, -sg_t2g), y = sg_t2g, fill = Player_initial_last)) +
  geom_col(show.legend = FALSE) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```
Here we can see the average shots to the green for the bottom 25 players 
The players on the right side have negative sq_t2g values causing them to loose 
strokes to the green, which is a key component of performance.

##top tournement & best 3 players by sg_t2g 

The 4 variables I created here are filtered to find the best 3 golfers in each 
of the top 4 tournaments in golf, also filtering by the year to keep the data 
as up-to-date as possible. 
```{r}
(top3_masters <- PGA %>% 
  filter(`tournament name` == 'Masters Tournament', season == 2022) %>%
  filter(!is.na(pos)) %>% 
  arrange(pos) %>% 
  slice(1, 3, 4))
  
```

```{r}
(top3_open <- PGA %>% 
  filter(`tournament name` == 'U.S. Open', season == 2021) %>%
  filter(!is.na(pos)) %>% 
  arrange(pos) %>% 
  slice(1:3))
```

```{r}
(top3_pgaChamp <- PGA %>% 
  filter(`tournament name` == 'PGA Championship', season == 2022) %>%
  filter(!is.na(pos)) %>% 
  arrange(pos) %>% 
  slice(1:3))
```

```{r}
(top3_playersChamp <- PGA %>% 
  filter(`tournament name` == 'The Players Championship', season == 2022) %>%
  filter(!is.na(pos)) %>% 
  arrange(pos) %>% 
  slice(1:3))
```

All 4 of these plots show the top 3 golfers in that specific tournament on the 
x-axis and on the y-axis, it shows their shot to the green. Then colored by their 
finishing position in the tournament.
```{r}
ggplot(top3_masters, aes(x = reorder(Player_initial_last, sg_t2g), y = sg_t2g, fill = factor(pos))) +
  geom_col()+
  labs(title = "Masters")
```
Here we can see that the top 3 golfers in the 2022 Masters have a high sg_t2g which 
makes sense and is what we are trying to prove.
```{r}
ggplot(top3_open, aes(x = reorder(Player_initial_last, sg_t2g), y = sg_t2g, fill = factor(pos))) +
  geom_col()+
   labs(title = "U.S Open")
```
Here we can see that the top 3 golfers in the 2022 U.S Open have a high sg_t2g which 
makes sense and is what we are trying to prove.

```{r}
ggplot(top3_pgaChamp, aes(x = reorder(Player_initial_last, sg_t2g), y = sg_t2g, fill = factor(pos))) +
  geom_col()+
   labs(title = "PGA Championship")
```
Here we can see that the top 3 golfers in the 2022 PGA Championship have a high sg_t2g which 
makes sense and is what we are trying to prove.
```{r}
ggplot(top3_playersChamp, aes(x = reorder(Player_initial_last, sg_t2g), y = sg_t2g, fill = factor(pos))) +
  geom_col()+
   labs(title = "The Players Championship")
```
Here we can see that the top 3 golfers in the 2022 Players Championship have a high sg_t2g which 
makes sense and is what we are trying to prove.





##corr
Let's find the correlation between performative stats and position, like sg_total and sq_t2g.

Taking only these 3 variables from PGA and setting them to PGA_performance
and taking out all the missing variables. 
```{r}
PGA_performance <- PGA[, c('sg_total', 'sg_t2g', 'pos'), use = "complete.obs"]
```



```{r}
(cor_matrix <- cor(PGA_performance, use = "complete.obs"))
```
We can see that sg_total and sg_t2g have a strong positive correlations (.78)
This means that players who gain more strokes overall also tend to gain more from
to too green.

sg_total and pos have a strong negative correlation(-.70). Which means that 
players with higher total scores tend to have lower positions. So obviously 
displaying better overall performance will lead to a better position in a tournament. 

sg_t2g and pos have a moderate negative correlation(-.55). Meaning that players 
Those are better from tee to green and usually finish better.

Here we can see that sg_total is super important in the finishing place, and sg_t2g 
is also a very important and key part of success in the game. 



```{r}
corrgram(PGA_performance)

```
*sg_total and sg_t2g have strong positive correlations 
*sg_total and pos have a strong negative correlation
*sg_t2g and pos have a moderate negative correlation

strokes gained to tee vs actual finish position




## PGA champ numbers




Finding all the data for the 2022 PGA Championship.
```{r}
(PGA_championship <- PGA %>%
  filter(`tournament name` == "PGA Championship", date == '2022-05-22') %>%
  distinct(Player_initial_last, strokes, pos, made_cut, sg_putt, sg_app, sg_arg, sg_ott, sg_t2g, sg_total, n_rounds))

```
```{r}
head(PGA_championship)
```

Finding the top 10 players in the tournament. 
```{r}
(PGA_top10chapm <- PGA_championship %>% 
  filter(!is.na(pos)) %>% 
  arrange(desc(pos)) %>% 
  filter(!is.na(pos)) %>% 
  arrange(pos) %>% 
  slice(1:10) %>% 
  dplyr::select(Player_initial_last))
```



Here I found all of the avg most important stats from the players performances. 
```{r}
(top_2022 <- PGA_championship %>%
group_by( Player_initial_last) %>%
  summarise(
    avg_sg_putt = mean(sg_putt, na.rm = TRUE),
    avg_sg_arg = mean(sg_arg, na.rm = TRUE),
    avg_sg_t2g = mean(sg_t2g, na.rm = TRUE),
    avg_total_sg = mean(sg_total, na.rm = TRUE)
  ) %>%
  arrange(desc(avg_total_sg)) %>%
  slice_head(n = 10))
```







Attempting to predict the winner and top 3 of the 2023 PGA Championship.
```{r}
recent_PGA <- PGA %>% 
  filter(season == 2022, made_cut == 1) %>%
  dplyr::select(Player_initial_last, pos, sg_putt, sg_app, sg_arg, sg_ott, sg_t2g, sg_total, `tournament name`)
```

```{r}
recent_PGA
```
```{r}
recent_PGA_clean <- recent_PGA %>%
  drop_na(sg_putt, sg_arg, sg_ott, sg_app, sg_t2g, sg_total, pos)
```

Creating a prediction modle using multiple regression.
```{r}
position_model <- lm(pos~ sg_putt + sg_arg + sg_ott +sg_app + sg_t2g + sg_total, data = recent_PGA_clean)
```

```{r}
summary(position_model)
```
Total strokes gained have a strong negative effect on position. Here, we can find that the best overall predictor is the player's overall strokes gained. This is especially true if their short game is solid. 


```{r}
predicted_positions <- predict(position_model)
```
```{r}
recent_PGA_wPrediction <- recent_PGA_clean %>% 
  mutate(predicted_pos = predicted_positions )
```


```{r}
ggplot(recent_PGA_wPrediction, aes(x = sg_t2g, y = pos))+
  geom_point()+
  geom_smooth()
```
This plot demonstrates that as players' strokes gained from tee to green
increase, their finishing position also tends to increase. Lower strokes gained 
from tee to green are better.


Here is where our new variable predicted_pos is showing us the top 10 players 
predicted to win the 2023 PGA Championship based on our model.
```{r}
top_predictions <-recent_PGA_wPrediction %>% 
  arrange(predicted_pos)

head(top_predictions, 10)
```


